---
title: "ClipMixcure Manual"
author: "Changchang Xu"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{ClipMixcure Manual}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  eval = FALSE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

# Overview

`ClipMixcure` supports multiply-imputed Weibull mixture cure modeling and combined likelihood profile (CLIP) inference. It is designed for settings where Rubin-style normal approximations may be unreliable and profile-based inference is preferred.

The main functions are:

- `mixcure.penal.mi()` for fitting mixture cure models across imputations
- `clip.mixcure()` for CLIP confidence intervals and p-values
- `pool.mixcure()` for Rubin-style pooled summaries
- `mixcure.clip.pdf()` for profile-density calculations used by CLIP
- `clipmixcure_fit()` as the adapter used by `Mixcuref.Meta`

# Package data

The package includes one multiply-imputation example dataset:

- `ANNbcBMdat5_miss`

This dataset contains missing values and is intended for use with `mice`.

# Multiple imputation workflow

```{r}
library(ClipMixcure)
library(mice)
library(survival)

data(ANNbcBMdat5_miss)

imp20 <- mice(ANNbcBMdat5_miss, m = 20, maxit = 20, seed = 24415, print = FALSE)
data20 <- complete(imp20, action = "long", include = TRUE)

data20$TN <- factor(with(data20,
  ifelse((Her2 == "1" & ER == 2 & PR == 2), "2", "1")
))

data20$LuminalA <- factor(with(data20,
  ifelse((Her2 == "1" & (ER == 1 | PR == 1) & KI67 == 1), "2", "1")
))

data20$LuminalB <- factor(with(data20,
  ifelse((Her2 == "1" & (ER == 1 | PR == 1) & KI67 == 2), "2", "1")
))

imp20sub <- as.mids(data20)
```

# Fitting across imputations

```{r}
mixcure_all20 <- with(
  data = imp20sub,
  expr = mixcure.penal.mi(
    Surv(Time, CENS == 1) ~ Her2 + LuminalA + TN + MENS0 + TUMCT,
    init = c(2, rep(0.1, 5), -10, rep(0.1, 5), 1),
    pl = FALSE
  )
)

class(mixcure_all20)
length(mixcure_all20$analyses)
```

# CLIP inference

```{r}
out_clip <- clip.mixcure(
  obj = mixcure_all20,
  pl = FALSE
)

out_clip
```

Firth-type penalized estimation can be requested by setting `pl = TRUE` in both the fitting and CLIP steps.

```{r}
mixcure_all20_penal <- with(
  data = imp20sub,
  expr = mixcure.penal.mi(
    Surv(Time, CENS == 1) ~ Her2 + LuminalA + TN + MENS0 + TUMCT,
    init = c(2, rep(0.1, 5), -10, rep(0.1, 5), 1),
    pl = TRUE
  )
)

out_clip_penal <- clip.mixcure(
  obj = mixcure_all20_penal,
  pl = TRUE
)
```

# Rubin-style pooling

```{r}
pooled <- pool.mixcure(mixcure_all20)
pooled
```

# Meta-layer adapter

The `clipmixcure_fit()` adapter is used by `Mixcuref.Meta` to route engine-specific requests. It accepts an `mc_spec` object and engine-specific control arguments.

```{r}
# spec is a model specification object created by Mixcuref.Meta
# fit <- clipmixcure_fit(spec, control = list(
#   method = "penal_mi",
#   formula = Surv(Time, CENS == 1) ~ Her2 + LuminalA + TN + MENS0 + TUMCT,
#   init = c(2, rep(0.1, 5), -10, rep(0.1, 5), 1),
#   pl = FALSE
# ))
```

# Notes

- `mixcure.penal.mi()` is intended to be used inside `with(mids, ...)`.
- `clip.mixcure()` consumes the resulting `mira` object.
- CLIP intervals are especially useful when normal approximations are questionable.
- Derived subtype variables such as `TN` and `LuminalA` should be created before converting back to a `mids` object with `as.mids()`.

# Session information

```{r, eval=T, echo=F}
pkgs <- c(
#  "ClipMixcure",
  "mice",
  "survival",
  "splines",
  "abind",
  "R.utils"
)

desc_field <- function(pkg, field) {
  x <- tryCatch(utils::packageDescription(pkg, fields = field), error = function(e) NA)
  if (is.null(x) || is.na(x)) "" else x
}

pkg_table <- data.frame(
  Package = pkgs,
  Version = sapply(pkgs, function(x) as.character(packageVersion(x))),
  Built   = sapply(pkgs, desc_field, field = "Built"),
  Remote  = sapply(pkgs, desc_field, field = "RemoteSha"),
  row.names = NULL,
  stringsAsFactors = FALSE
)

knitr::kable(pkg_table, caption = "Package Versions")

cat("\n\n")

#sessionInfo()

```
